---
title: "ChimpSniffingCleaning"
author: "Noah Stetson"
format: html
editor: visual
---

### Uploading stuff

Load library

```{r}
library(tidyverse)
library(dplyr)
```

Load data

```{r}
sniffing <- read.csv("https://raw.githubusercontent.com/nwstetson/ChimpSniffing/main/Data/sniffing_allData_2014-2021.csv")

biomarker <- read.csv("https://raw.githubusercontent.com/nwstetson/ChimpSniffing/main/Data/demographyForNetworks_14Aug2019.csv")
```

### Tidying data

Take out commas after names

```{r}
sniffing$Individual <- str_replace(sniffing$Individual, ",", "")
sniffing$ID2 <- str_replace(sniffing$ID2, ",", "")
```

change "IndividualID" in biomarker df to be called "Individual" so it matches sniffing df

```{r}
colnames(biomarker) [colnames(biomarker) == "individualID"] <- "Individual"
```

Add sex and DOB from biomarker into sniffing based on individual

```{r}
sniffing <- inner_join(sniffing, biomarker, by = "Individual")
sniffing <- sniffing[, !(colnames(sniffing) %in% c("father","mother","Maturity","ID_notes","neighborhood","Focal"))]
sniffing <- sniffing %>% relocate (DOB, Sex, .before = Behavior)
```

#### Use DOB and Date to find Age and make Age a column

Convert DOB to be in same format as data

```{r}
sniffing$DOB <- as.Date(as.character(sniffing$DOB))
sniffing$Date  <- as.Date(as.character(sniffing$Date), format = "%m/%d/%y")
```

Calculate age per sample and put age next to individual

```{r}
sniffing$Age <- round(as.numeric((sniffing$Date - sniffing$DOB) / 365.25), 2)
sniffing <- sniffing %>% relocate (Age, .before = Sex)
```

Remove line that says to remove it (lines 189 and 190)

```{r}
sniffing <- sniffing[-c(189:190), ]

sniffing <- sniffing[, !(colnames(sniffing) %in% c("Delete"))]
```

Now need to summarize data into a table and group by individual and calculate the mean age and total number of sniffs

Columns will be: Individual, Mean Age, Sex, Total Sniffs

```{r}
# remane "sniff and sniff.ground" to 1 bc it's one count of a sniff
sniffing$Behavior <- str_replace(sniffing$Behavior, "Sniff.Ground", "1")
sniffing$Behavior <- str_replace(sniffing$Behavior, "Sniff", "1")
# rename "behavior" column to be called "sniffs"
sniffing <- sniffing %>% rename(Sniff = Behavior)
```

What if I just saved it as a new file and then did the summary on that

```{r}
write.csv(sniffing, "./data/sniffing_clean.csv", row.names=FALSE)

sniffing_clean <- read.csv("https://raw.githubusercontent.com/nwstetson/ChimpSniffing/main/Data/sniffing_clean.csv")
```

```{r}
sniff_summary <- sniffing_clean %>%
  group_by(Individual) %>%
  summarize(Age = mean(Age, na.rm = T), Sex = first(Sex), TotalSniffs = sum(Sniff, na.rm = T))
```

wtf it worked

slay

Okay now I'm going to make a visualization with it

```{r}
write.csv(sniff_summary, "./data/sniff_summary.csv", row.names=FALSE)
```
